---
title: "Homework_3"
author: "Brennan Baker"
date: "October 9, 2018"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem 1

Load packages
```{r packages}
library(tidyverse)
library(p8105.datasets)
```



*Tidy the data*


```{r tidy brfss data}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  separate(locationdesc, into = c("state", "county"), sep = "- ") %>% 
  select(year, state, county, response, data_value) %>% 
  mutate(response = tolower(response)) %>% 
  mutate(response = ordered(response, levels = c("excellent", "very good", "good", "fair", "poor")))
```

I cleaned the data by: cleaning the names; keeping only the rows where the topic is overall health; separating "locationdesc" into state and county columns; and excluding unnecessary variables. I converted response to a factor varaible and arranged it with "Excellent"" on top.

*In 2002, which states were observed at 7 locations?*

```{r}
brfss %>% 
  filter(year == "2002") %>% 
  distinct(state, county, .keep_all = TRUE) %>% 
  count(state) %>% 
  filter(n == "7") %>% 
  nrow()
```

There were 3 states observed at 7 locations. 

*Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010*

```{r}
brfss %>%
  group_by(year) %>% 
  count(state) %>% 
    ggplot(aes(x = year, y = n, color = state)) +
  geom_line() +
  theme_bw()
```

*Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.*


```{r}
brfss %>%
  filter(year %in% c("2002", "2006", "2010")) %>% # Only need these years
  filter(str_detect(state, "NY")) %>%  # Only need NY
  spread(key = response, value = data_value) %>% # spread and mutate to create new col containing the proportion of excellent responses
  janitor::clean_names() %>% 
  mutate(proportion_excellent = (excellent)/(fair + good + poor + excellent + very_good)) %>%
  group_by(year) %>% 
  filter(!is.na(proportion_excellent)) %>%
  summarize(mean = mean(proportion_excellent), std_dev = sd(proportion_excellent)) %>% 
  knitr::kable(digits = 4)
```


*For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.*
